Stack Safety as a Security Property

### 2020-07-09 main.tex
typo
A pass on related work

CHERI \rb{cite}

\citet{Skorstengaard+19b} introduce a calling convention to preserve local state
encapsulation and well-bracketed control flow by the use of local capabilities.
Although the required hardware support is readily available in capability
machines like CHERI, this technique incurs potentially prohibitive performance
costs, because it requires the unused part of the stack to be cleared whenever a
security boundary is crossed. In the discussion, ``while [the authors] claim
that [their] calling convention enforces control-flow correctness, [they] do not
prove a general theorem that shows this, because it is not clear what such a
theorem should look like,'' noting that the correctness property enjoyed by
their technique ``is not made very explicit.''

StkTokens \citep{Skorstengaard+19} protects the stack by enforcing local state
encapsulation and well-bracketed control flow. It does so by defining a calling
convention that makes use of linear capabilities for stack and return pointers
operating on a single shared stack, and by requiring that protected components
avoid compromising their own security by following certain simple rules.
Informally, local state encapsulation is defined by restricting accesses to the
range of memory allocated to the current stack frame, and well-bracketed control
flow by only allowing returns from the topmost frame to the immediately adjacent
frame below. Formally, it improves on \citet{Skorstengaard+19b} by building
those properties into the semantics of a capability machine with a built-in call
stack and call and return instructions, which is proven fully abstract with
respect to a more concrete capability machine that replaces those
pseudo-instructions with their calling convention.

\rb{As far as I can tell, while their notion of components and interfaces seems
  standard, observable behavior appears to be limited to cotermination. The
  previous paper makes explicit that (at least at that point) they are not
  interested in externally observable side effects like console output or memory
  access traces.}

We are confident \rb{how strong a claim to make, with substantiation} that
StkTokens satisfies the eager stack safety property, and even the stronger eager
testing property \rb{refs}. The main limitation of StkTokens is its reliance on
linear capabilities, as it is unclear now they could be added to practical
capability machines, especially in terms of efficiency.
%
\rb{On the other hand, we know that these can be implemented as micropolicies,
  cf. linear return capabilities in yannis-report (what about performance?)}

\rb{There are quite a few parallels at various levels, although the definitions
  are overall more implicit and not given formally.}

@@

  We have presented a pure notion of stack safety that omits many complicating
  factors common to real system. In particular, our model as presented exploits
  the simplifying assumption that privilege never increases during a function
  call, and that therefore a contour computed once on the entry to a function
  describes its privilege throughout. Our future plans are to extend the model
  with common language features, some of which violate this assumption. Here we
  describe the basic principles by which our model can be extended to some of
  these cases.

>>>>

%% CHERI \rb{cite}

\paragraph{Formal Stack Safety on Capability Machines}
%
To our knowledge, the first and only line of work that has attempted to furnish
a positive and formal characterization of the meaning of stack safety was
inaugurated by \citet{Skorstengaard+19b}, who introduce a calling convention
that uses local capabilities to preserve local state encapsulation and
well-bracketed control flow, using a logical relation to reason about the stack
safety of concrete programs. Although the required hardware support is readily
available in capability machines like CHERI, this technique incurs potentially
prohibitive performance costs, because it requires the entire unused part of the
stack to be cleared whenever a security boundary is crossed. Their logical
relation captures capability safety without ``externally observable side-effects
(like console output or memory access traces).'' In the discussion, ``while [the
authors] claim that [their] calling convention enforces control-flow
correctness, [they] do not prove a general theorem that shows this, because it
is not clear what such a theorem should look like,'' noting that the correctness
property enjoyed by their technique ``is not made very explicit.''

StkTokens \citep{Skorstengaard+19} is a natural continuation of that work. Like
its predecessor, it aims to protect the stack by enforcing local state
encapsulation and well-bracketed control flow. It does so by defining a new
calling convention that makes use of linear capabilities for stack and return
pointers. The convention operates on a single shared stack and requires that
protected components avoid compromising their own security by following certain
simple rules---like avoiding to leak their private capability seals. More
precisely but still informally, local state encapsulation is defined as
restricting accesses to the range of memory allocated to the current stack
frame, and well-bracketed control flow as only allowing returns from the topmost
frame to the immediately adjacent frame below. Formally, it improves on
\citet{Skorstengaard+19b} by building those properties into the semantics of a
capability machine with a built-in call stack and call and return instructions,
which is proven fully abstract with respect to a more concrete capability
machine that replaces those pseudo-instructions with their calling convention.
Their proof of full abstraction uses a standard notion of components, which
import and export functions through their interfaces; their model of
observations is limited to cotermination.

We are confident that the linear capability machines introduced by
\citep{Skorstengaard+19}, in combination with the StkTokens calling convention,
can be modeled in our framework and satisfy our definition of stack safety in
\cref{sec:lse-and-wbcf}. Further, we expect them to satisfy the even stronger
property developed in \cref{sec:enforcement}. Proving these conjectures would
involve materializing their definitions of local state encapsulation and
well-bracketed control flow, which exist only in enmeshed form as part of the
semantics of the capability machines. The main practical limitation of StkTokens
is its reliance on linear capabilities, as it is unclear now they could be added
to practical capability machines, especially in terms of efficiency---previous
work on micropolicies \citep{yannis-report} has shown how to use that framework
to implement linear return capabilities.

\paragraph{Protecting the Stack with Micropolicies}
%
\citet{DBLP:conf/sp/RoesslerD18} consider a standard attacker model where all
attacks against stack data are in scope (but not side channels or hardware
attacks), and study the protection of stack data through three families of
micropolicies that tag stack objects with a pair of frame and object identifiers
used to validate accesses to the stack: Return Address Protection (which
prevents an adversary from overwriting designated return addresses), Static
Authorities (which only allows the code of a function to access the stack frames
of its own dynamic instances), and Depth Isolation (described in
\cref{sec:enforcement}). All these policies exploit various kinds of spatial and
temporal locality of stack memory and local call graphs, as well as information
generated by the compilation toolchain, to strike various balances between
precision and cacheability, evaluated through benchmarks that demonstrate
limited performance overhead. In addition to the baseline ``eager'' policies,
they propose a number of lazy optimizations, notably Lazy Tagging and Lazy
Clearing, discussed in \cref{sec:lazy}, with associated improvements in
performance.

@@

  We have presented a pure notion of stack safety that omits many complicating
  factors common to real systems. In particular, our model as presented exploits
  the simplifying assumption that privilege never increases during a function
  call, and that therefore a contour computed once on the entry to a function
  describes its privilege throughout. Our future plans are to extend the model
  with common language features, some of which violate this assumption. Here we
  describe the basic principles by which our model can be extended to some of
  these cases.
